---
title: "Lab 5"
output: html_document
date: "2024-03-29"
---

```{r}
library(tidyverse)
library(ggplot2)
load("~/Desktop/RWork/V0500_Lecture1”/BRFSS2022_rev.RData")


# first want to make some factors into numbers
brfss22$Age_midpt <- fct_recode(brfss22$X_AGEG5YR, "21" = "Age 18 to 24",
                                "27" = "Age 25 to 29", "32" = "Age 30 to 34",
                                "37" = "Age 35 to 39", "42" = "Age 40 to 44",
                                "47" = "Age 45 to 49", "52" = "Age 50 to 54",
                                "57" = "Age 55 to 59", "62" = "Age 60 to 64",
                                "67" = "Age 65 to 69", "72" = "Age 70 to 74",
                                "77" = "Age 75 to 79", "82" = "Age 80 or older",
                                NULL = "Dont know/Refused/Missing")
brfss22$Age_midpt <- as.numeric(levels(brfss22$Age_midpt))[brfss22$Age_midpt]
brfss22$income_midpoint <- fct_recode(brfss22$INCOME3, 
                                      "7500" = "Household income less than $10,000",
                                      "12500" = "Less than $15,000 ($10,000 to less than $15,000)",
                                      "17500" = "Less than $20,000 ($15,000 to less than $20,000) ",
                                      "22500" = "Less than $25,000 ($20,000 to less than $25,000) ",
                                      "30000" = "Less than $35,000 ($25,000 to less than $35,000) ",
                                      "42500" = "Less than $50,000 ($35,000 to less than $50,000) ",
                                      "62500" = "Less than $75,000 ($50,000 to less than $75,000)",
                                      "87500" = "Less than $100,000 ($75,000 to less than $100,000)",
                                      "125000" = "Less than $150,000 ($100,000 to less than $150,000)",
                                      "175000" = "Less than $200,000 ($150,000 to less than $200,000)",
                                      "210000" = "$200,000 or more",
                                      NULL = "Dont know/Not sure",
                                      NULL = "Refused")
brfss22$income_midpoint <- as.numeric(levels(brfss22$income_midpoint))[brfss22$income_midpoint]



brfss22$Educ_number <- fct_recode(brfss22$EDUCA, 
                                  "0" = "Never attended school or only kindergarten", 
                                  "4.5" = "Grades 1 through 8 (Elementary)",
                                  "10" = "Grades 9 through 11 (Some high school)",
                                  "12" = "Grade 12 or GED (High school graduate)",
                                  "14" = "College 1 year to 3 years (Some college or technical school)",
                                  "16" = "College 4 years or more (College graduate)",
                                  NULL = "Refused" )
brfss22$Educ_number <- as.numeric(levels(brfss22$Educ_number))[brfss22$Educ_number]

```


```{r}
# this level name can create problems later so fix it here
levesbrfssmarijanXPRACE27 <- "dont know not sure"

select1 <- !is.na(brfss22$MARIJAN1)
brfss_marijan <- subset(brfss22, select1 )


p_cannabis_age <- ggplot(data = brfss_marijan,
                         mapping = aes(x = Age_midpt,
                                       y = MARIJAN1))
p_cannabis_age + geom_smooth()

model_1 <- lm(MARIJAN1 ~ Age_midpt, data = brfss_marijan)
summary(model_1)
```


```{r}
#According to the data, race is significant because depending on the participants race there is a decrease/increase in marijuana usage.
#Based on the data, if you identify as Black or African American there is a significantly less usage of marijuana compared to if you are
#American Indian or Alaskan Native. Another significant correlation is Marijuana use and eduation. There is a significant correlation between 
#those that went to some high school and marijuana usage. 
model_2 <- lm(MARIJAN1 ~ Age_midpt + X_PRACE2 + X_HISPANC + EDUCA, data = brfss_marijan)
summary(model_2)



# maybe get fancy
require(stargazer)
stargazer(model_2, type = "?")
# play with stargazer instead of summary, it can look nicer!
# it might throw an error at top but then work fine, just ignore it

summary(brfss_marijan)

#Some additional measures that I incorporated for the 3rd model are acetouch and acehurt. A good predictor of marijuana usage 
#is if the participant was touced once or more than once. Another good predictor of marijuana usage is if the participant 
#was hurt once or more than once. These are good predictors of marijuana usage because their interaction with marijuana usage was 
#very significant compared to those that said don't know or refused to answer. 

model_3 <- lm(MARIJAN1 ~ Age_midpt + ACETOUCH + ACEHURT1, data = brfss_marijan)
summary(model_3)

#No depression means less marijuana usage (negative sign)
#In this model I am looking at marijuana, depression, and income. In this model those that 
#answered no to depression, answered don't know, or refused to answer have less marijuana usage.
#In addition, there was a significant link between income and less marijuana usage. 
model_4 <- lm(MARIJAN1 ~ ACEDEPRS + income_midpoint, data= brfss_marijan)
summary(model_4)

#graph
levels(brfss22$ACETOUCH)

brfss_marijan$brfss_marijan_touch <- fct_recode(brfss_marijan$ACETOUCH, "0" = "Adverse Childhood Exper, never:  How often did anyone at least 5 years older than you or an adult, ever touch you sexually",
                                                          "1"= "once",
                                                          "1"= "more than once",
                                                          NULL = "dont know not sure",
                                                          NULL = "refused")
brfss_marijan$brfss_marijan_touch <- as.numeric(levels(brfss_marijan$brfss_marijan_touch))[brfss_marijan$brfss_marijan_touch]

summary(brfss_marijan$brfss_marijan_touch)
                                                          
model_4_graph <- ggplot(data = brfss_marijan,
                         mapping = aes(x = MARIJAN1,
                                       y = brfss_marijan_touch))
model_4_graph + geom_point(alpha=0.05)+
  geom_smooth()+
labs(x = "Marijuana", y = "Touched",
     tile = "Marijuana usage in those that have been touched", 
     caption = "Source: brfss_marijan.")

```Compared to model 4, the variables in model three increased the r squared which means that they really help predict. These variables are Age midpoint, acetouch, and acehurt. I was surprised that the r squared didn't increase as much in model 4 because one of the variables in model 4 is depression and one would think that it has a significant correlation with marijuana usage. 


